Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-09-03 vesicle_code/docs/permeableVes/main.tex
worked on pnas paper

We further analyze the transient dynamics by considering the vesicle
velocity~\eqref{eqn:vesVelocity} for small $\beta$. Two terms contribute
to the vesicle velocity: (i) the permeability velocity, $\beta(\ff \cdot
\nn)\nn$, and (ii) the force balance velocity, $\SS[\ff]$. At early
times, since $\beta \ll 1$, the permeability velocity is much smaller
than the force balance velocity and the vesicle dynamics resembles that
of a relaxing impermeable vesicle. As the force balance velocity
decreases, the permeability velocity becomes significant, and the
vesicle begins to inflate to a circle. This results in the membrane
force and velocity both converging to zero. Letting $\beta = 10^{-3}$,
the force balance velocity (blue) and the permeability velocity (red)
are plotted at three times in Figure~\ref{fig:vesVelocity}. We also plot
the velocity of an identically initialized impermeable vesicle (green).
At the early time $t=0.5$, the permeability velocity is small relative
to the force balance velocity, and the force balance velocity matches
with the impermeable vesicle velocity. At time intermediate time $t=5$,
all three velocities are comparable in size, and the semi-permeable
vesicle begins to inflate. Finally, at time $t=50$, the impermeable
vesicle is at its steady state, while the semi-permeable vesicle is
still inflating. The bottom right plot of Figure~\ref{fig:vesVelocity}
is the maximum norm of the different velocity contributions as a
function of time, and the black vertical lines correspond to the times
reported in the other three subplots.

%  \input{figures/ellipseVelocity1.tikz}
%  \input{figures/ellipseVelocity2.tikz}
%  \input{figures/ellipseVelocity3.tikz}
%  \input{figures/ellipseVelocityNorms.tikz}
\begin{figure}[htp]
  \centering
%  \input{figures/ellipseVelocity1.tikz}
%  \input{figures/ellipseVelocity2.tikz}
%  \input{figures/ellipseVelocity3.tikz}
%  \input{figures/ellipseVelocityNorms.tikz}
  \includegraphics[width=0.48\linewidth]{figures/ellipseVelocity1.pdf}
  \includegraphics[width=0.48\linewidth]{figures/ellipseVelocity2.pdf}
  \includegraphics[width=0.48\linewidth]{figures/ellipseVelocity3.pdf}
  \includegraphics[width=0.48\linewidth]{figures/ellipseVelocityNorms.pdf}
  \caption{\label{fig:vesVelocity} The velocity contributions due to
  force balance (blue) and permeability (red). The green dashed line is
  the velocity of an impermeable vesicle with the same initial shape as
  the semi-permeable vesicle. The bottom right plot is the max norm of
  each contributor to the vesicle velocities. The three black lines
  correspond to the times reported in the other three plots.}
\end{figure}

We next investigate the interplay between the tension, transmembrane
flux, and curvature of a semi-permeable vesicle. We let $\beta = 1$ and
initialize the vesicle with regions of both positive and negative
curvature. The vesicle tension, $\Lambda$, and the flux, $\ff \cdot
\nn$, are plotted in Figure~\ref{fig:starTensionFlux}. At early times,
regions of large negative curvature have negative tension and regions
with large positive curvature have positive tension. In addition,
regions with large curvature have a large inflow of fluid (positive
flux), while the flatter regions have an outflow of fluid. At these
early times, the net effect is vesicle deflation. As the vesicle starts
to smooth out, the same general tension and flux profiles persist, but
the net flux is positive and the vesicle inflates. At later times, the
tension starts to become positive and the vesicle continues to inflate
towards a circle. Finally, the tension transitions to a positive
constant function, the membrane force and flux approach zero, and the
vesicle shape becomes positive.

>>>>

We further analyze the transient dynamics by considering the vesicle
velocity~\eqref{eqn:vesVelocity} for small $\beta$. Two terms contribute
to the vesicle velocity: (i) the permeability velocity, $\beta(\ff \cdot
\nn)\nn$, and (ii) the force balance velocity, $\SS[\ff]$. Since $\beta
\ll 1$, the permeability velocity is much smaller than the force balance
velocity at early times, and the semi-permeable vesicle dynamics
resembles those of an impermeable vesicle. As the force balance velocity
decreases, the permeability velocity becomes significant, and the
vesicle begins to inflate to a circle. This results in the membrane
force and velocity both converging to zero. Letting $\beta = 10^{-3}$,
the normal component of the force balance velocity (blue) and the
permeability velocity (red) are plotted at three times in
Figure~\ref{fig:vesVelocity}. We also plot the velocity (green) of an
identically initialized impermeable vesicle. At the early time $t=0.5$,
the permeability velocity is negligible relative to the force balance
velocity, and the force balance velocity matches with the impermeable
vesicle velocity. At intermediate time $t=5$, all three velocities are
comparable in size, and the semi-permeable vesicle begins to inflate.
Finally, at time $t=50$, the impermeable vesicle is near steady state,
while the semi-permeable vesicle is still inflating. The right-most
plot of Figure~\ref{fig:vesVelocity} is the maximum norm of the
different velocity contributions as a function of time, and the black
vertical lines correspond to the times reported in the other three
subplots.

\begin{figure*}[htp]
  \centering
%  \input{figures/ellipseVelocity1.tikz}
%  \input{figures/ellipseVelocity2.tikz}
%  \input{figures/ellipseVelocity3.tikz}
%  \input{figures/ellipseVelocityNorms.tikz}
  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity1.pdf}
  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity2.pdf}
  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocity3.pdf}
  \includegraphics[width=0.24\linewidth]{figures/ellipseVelocityNorms.pdf}
  \caption{\label{fig:vesVelocity} The velocity contributions due to
  force balance (blue) and permeability (red). The green line is the
  velocity of an impermeable vesicle with the same initial shape as the
  semi-permeable vesicle. The right-most plot is the maximum norm of
  each velocity contribution. The black lines correspond to the times
  reported in the other three plots.}
\end{figure*}

We next investigate the interplay between the tension, transmembrane
flux, and curvature of a semi-permeable vesicle. We let $\beta = 1$ and
initialize the vesicle with regions of both positive and negative
curvature. The vesicle tension, $\Lambda$, and the flux, $\ff \cdot
\nn$, are plotted in Figure~\ref{fig:starTensionFlux}. At early times,
regions of large negative curvature have negative tension and regions
with large positive curvature have positive tension, and the same result
holds for an impermeable vesicle. In addition, regions with large
curvature have a large inflow of fluid (positive flux), while the
flatter regions have an outflow of fluid. At these early times, the net
effect is vesicle deflation. As the vesicle shape relaxes, the same
general tension and flux trends persist, but the net flux is positive
and the vesicle inflates. At later times, the tension becomes positive
and the vesicle continues to inflate towards a circle. Finally, the
tension transitions to a positive constant function, the membrane force
and flux approach zero, and the vesicle shape reaches its steady
circular shape.
